---
title: "Peer-Graded Point-Referenced Spatial Data and Gaussian Processes"
output: html_document
date: "2023-07-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(mapproj)
library(geoR)
library(leaflet)
library(gridExtra)
library(stringr)

rain <- readr::read_table('https://raw.githubusercontent.com/DapperDataAnalyst/Advanced_Predictive_Models/main/rain.txt')

```

### Part (a)
```{r}
# Convert altitude from feet to miles
# There are 5,280 feet per mile
rain$altitude <- rain$altitude * (1/5280)

# Plot rainfall and altitude
rain_alt_plot <- ggplot(data = rain, aes(x = x, y = y)) + 
   geom_point(aes(size = rainfall, color = altitude)) +
  ggtitle('Rainfall & Altitude by Longitude & Latitude') +
  theme_bw()


rain_alt_plot

```

There does appear to be spatial dependence between altitude and rainfall amounts. This can be seen most clearly in the plot around x = 300, where the points are small in diameter and light in color. This indicates a small rainfall amount and a high altitude, respectively. From this exploratory analysis, we may form the preliminary impression that high altitudes see rather little rainfall compared to lower altitudes.


### Part (b)
```{r}
# Fit a linear regression model using lm()
model <- lm(sqrt(rainfall) ~ altitude, data = rain)

# Extract the estimated parameter values
coefficients <- coef(model)

# Extract the R-squared value
r_squared <- summary(model)$r.squared

# Extract the estimated error variance
error_variance <- summary(model)$sigma^2

coefficients
r_squared
error_variance

```
The estimated regression function in the form $y = \beta_0 + \beta_1 x + \epsilon$ is as follows:
$$y = `r coefficients[1]` + (`r coefficients[2]`) x + \epsilon$$

The estimated error variance is `r error_variance`

The model yields an $R^2$ of `r r_squared`, indicating that `r r_squared * 100` percent of the variation in the square root of rainfall is explained by altitude.


### Part (c)
```{r}
# Find distances between points in each pair
distances <- dist(rain[,1:2])

# Convert distances to a data frame so it can be fed to ggplot
dist_df <- data.frame(Distance = as.vector(distances))

# Make histogram
histo <- ggplot(dist_df, aes(x = Distance, y = ..count../sum(..count..))) +
  geom_histogram(binwidth = 20,
                 center = 10,
                 fill = '#009E73', color = 'black') +
  scale_x_continuous(breaks = seq(from = 0, to = 300, by = 20),
                     labels = seq(from = 0, to = 300, by = 20)) +
  labs(x = "Distance", y = "Frequency", title = "Euclidean Distance Histogram") +
  theme_bw()

histo

```

### Part (d)
```{r message=FALSE, warning=FALSE}
rain$resid <- model$residuals

dist_matrix <- as.matrix(distances)
dist_matrix[upper.tri(dist_matrix)] <- 0


# Get all combinations of elements from vec1 and vec2
combinations <- as.data.frame(which(dist_matrix != 0, arr.ind = T))

combinations <- cbind(combinations,dist_df)

combinations$resid_row <- rain$resid[combinations$row]
combinations$resid_col <- rain$resid[combinations$col]
combinations$category <- cut(combinations$Distance,
                             breaks = seq(from = 0, to = 300, by = 20),
                             labels = c("[0,20)", "[20,40)", "[40,60)", "[60,80)", "[80,100)", "[100,120)", "[120,140)", "[140,160)", "[160,180)", "[180,200)", "[200,220)", "[220,240)", "[240,260)", "[260,280)", "[280,300)"),
                             include.lowest = TRUE)

combinations$bin_center <- floor(combinations$Distance / 20) * 20 + 20/2

correlations <- combinations %>%
  group_by(category) %>%
  summarise(correlation = cor(resid_row, resid_col), count = n())

bins_and_centers <- combinations %>% select(category, bin_center)

correlations <- correlations %>%
  left_join(bins_and_centers) %>% unique()

ggCorr <- ggplot(correlations, aes(x = bin_center, y = correlation)) +
  geom_point(aes(size = count, color = count)) +
  labs(x = "Bin Center", y = "Correlation", title = "Correlation Coefficient by Bin Center") +
  scale_color_viridis_c() +
  theme_bw()

ggCorr


```


### Part (e)
Because we are dealing in only two dimensions. In two dimensions, a maximum of three points can be placed equidistant to one another, at the vertices of an equilateral triangle. Maintaining equidistance with more than three points requires more than two dimensions. In two dimensions, adding more than three points requires that some pairwise distances be shorter than others, and that the shorter distances be more numerous than the longer distances.


### Part (f)
```{r message=FALSE, warning=FALSE}
rain_geo <- as.geodata(
  cbind(rain$x,
        rain$y,
        sqrt(rain$rainfall),
        rain$altitude),
  coords.col = 1:2,
  data.col = 3,
  covar.col = 4
)

fit <- likfit(
  rain_geo,
  cov.model = "exponential",
  ini.cov.pars = c(5, .01),
  fix.nugget=T, 
  nugget=0,
  trend = as.formula(~covar1,rain_geo))

fit$parameters.summary
```

From the outputted table, we find that $\beta_0 = `r fit$parameters.summary[1,2]`$, $\beta_1 = `r fit$parameters.summary[2,2]`$, $\sigma^2 = `r fit$parameters.summary[4,2]`$, and $\phi = `r fit$parameters.summary[5,2]`$


### Part (g)
```{r message=FALSE, warning=FALSE}
likfit_resids <- fit$model.components$residuals
combinations$likfit_resid_row <- likfit_resids[combinations$row]
combinations$likfit_resid_col <- likfit_resids[combinations$col]

likfit_correlations <- combinations %>%
  group_by(category) %>%
  summarise(correlation = cor(likfit_resid_row, likfit_resid_col), count = n())

likfit_correlations <- likfit_correlations %>%
  left_join(bins_and_centers) %>% unique()

ggCorr + geom_line(data = likfit_correlations, aes(x = bin_center, y = correlation),
                   linewidth = 1,
                   color = '#AAAAAA') +
  ggtitle('Correlation Coefficient & LikFit Correlation by Bin Center')

```


### Part (h)
```{r message=FALSE, warning=FALSE}
summary(fit)
```

From this summary of the models, we find that the non-spatial AIC is $578.3$ and the spatial AIC is $503.5$

A lower AIC value indicates better model fit, and so the spatial model is a better fit to the data than the non-spatial.


### Part (i)
We would also require information on an unmonitored location's altitude, as rainfall is correlated to altitude. In the kriging demo, there was no covariate for scallop catch size analogous to what we have here with altitude being a covariate to rainfall amount.


