---
title: "Staff-Graded Time Series EDA and Classical Models"
output: html_document
date: "2023-06-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(gridExtra)
library(tidyverse)
library(astsa)
library(xts)
library(lubridate)
library(stats)

temps <- readr::read_csv('https://raw.githubusercontent.com/DapperDataAnalyst/Advanced_Predictive_Models/main/MeanDallasTemps.csv')
```

## Question 3
### Part (a)

```{r include = FALSE}
S <- 12
```
Because the data are monthly, a seasonal lag term of $S=`r S`$ would be most appropriate, as we would expect a given month in one year to have average temperatures very similar to the same month in another year.

### Part (b)
```{r message=FALSE, warning=FALSE}

arg_names <- c('p','d','q','P','D','Q')

arg_count <- length(arg_names)
l <- rep(list(0:1), arg_count)

# Make data frame containing all combinations of the parameters
combo_frame <- data.frame(expand.grid(l))

# Clean up column names
colnames(combo_frame) <- arg_names

# Remove rows as instructed in homework document
rows_to_remove <- combo_frame[combo_frame$P==1 & combo_frame$D==0 & combo_frame$Q==1 & combo_frame$d==0,]
combo_frame <- combo_frame %>% anti_join(rows_to_remove)


# Make custom function to run SARIMA on all parameter combos
mass_sarima <- function(vect){
 fit <- sarima(temps$AvgTemp,
  p = vect[1],
  d = vect[2],
  q = vect[3],
  P = vect[4],
  D = vect[5],
  Q = vect[6],
  S = S,
  no.constant = TRUE,
  details = FALSE)
 return(fit$AICc)
}

# Bind AICc outputs to combo_frame
combo_frame <- cbind(combo_frame,apply(combo_frame, MARGIN = 1, mass_sarima))

# Clean up column name
names(combo_frame)[7] <- 'AICc'

# Find lowest AICc and return the parameter set that generates it
best_AICc <- min(combo_frame[,7])
best_parameters <- combo_frame[which.min(combo_frame$AICc),]
best_parameters

```

From this analysis, we find that the minimum AICc value is `r best_parameters[,'AICc']`. This value is found at parameter values p=`r best_parameters[,'p']`, d=`r best_parameters[,'d']`, q=`r best_parameters[,'q']`, P=`r best_parameters[,'P']`, D=`r best_parameters[,'D']`, Q=`r best_parameters[,'Q']`.

### Part (c)
```{r 3Qc}

# Define ideal parameters from Part (b)
p <- best_parameters[,'p']
d <- best_parameters[,'d']
q <- best_parameters[,'q']
P <- best_parameters[,'P']
D <- best_parameters[,'D']
Q <- best_parameters[,'Q']
forecasted_temps <- data.frame()


# Find row of Jan 2015
last_real_row <- which(temps$Month == '2015-01-01')


for (i in 1:(nrow(temps) - last_real_row - 11)) {
  
  
  # Slice temps data to only include training data
  temp_train <- temps[1:last_real_row,]
  
  # Run SARIMA forecast
  forecast_model <- sarima.for(temp_train$AvgTemp, 
                        n.ahead = S, 
                        p = p, 
                        d = d, 
                        q = q,
                        P = P,
                        D = D,
                        Q = Q,
                        S = S,
                        plot = F)
  
  # Collect forecast for one year from now and add to forecasted_temps
  next_year_forecast <- forecast_model$pred[12]
  forecast_row <- c(as.character(temps$Month[last_real_row+12]), as.numeric(next_year_forecast))
  forecasted_temps <- rbind(forecasted_temps,forecast_row)

  last_real_row <- last_real_row+1
  
}

# Clean up column headers
colnames(forecasted_temps) <- c('Date', 'Temp')







```












