---
title: "Staff-Graded: Spatial Point Patterns and Processes"
output: html_document
date: "2023-06-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(spatstat)
library(sf)
library(sp)
library(rgdal)
library(ggmap)

latlong <- readr::read_csv("https://raw.githubusercontent.com/DapperDataAnalyst/Advanced_Predictive_Models/main/chi_markets.csv")
load("chi_supermarkets_ppp.Rdata")
```

### Part (a)
```{r message=FALSE, warning=FALSE}
# Create a SpatialPoints object with latitude and longitude coordinates
points <- SpatialPoints(latlong, proj4string = CRS("+proj=longlat +datum=WGS84"))

# Define UTM coord system
utm_crs <- CRS("+proj=utm +zone=16 +datum=WGS84")

# Convert latlong to UTM
points_utm <- spTransform(points, utm_crs)

# Pull out x and y components and put in clean dataframe
utm_x <- coordinates(points_utm)[, 1]
utm_y <- coordinates(points_utm)[, 2]
utm_coords <- data.frame(
  x = utm_x,
  y = utm_y
)

# Go back to using lat/long coords
# Specify the lat/long coordinates of the map area
map_area <- c(left = min(latlong$lon), bottom = min(latlong$lat), right = max(latlong$lon), top = max(latlong$lat))
map <- get_stamenmap(bbox = map_area, maptype = "toner-lite")

# Plot
ggmap(map) +
  geom_point(data = latlong, aes(x = lon, y = lat), color = "red", size = 1) +
  xlab(NULL) +
  ylab(NULL) +
  ggtitle("Market locations in Chicago area")
```

The top right corner of the rectangle is open water, not buildable land.


### Part (b)
```{r}
plot(markets_ppp, main = "Markets in the Chicago area")
```

This spatial pattern appears to exhibit clustering, as many markets are located near the eastern border of the defined region, especially in the northern half of the defined region. Additionally, the intensity of the spatial events does not appear to be constant over time.


### Part (c)
```{r message=FALSE, warning=FALSE}
model <- ppm(markets_ppp, ~ polynom(x, y, 2))

plot(model,
  how = "image",
  se = FALSE,
  col = grey(seq(1, 0, length = 100))
)


model$coef

intensity_point <- predict(model, locations = data.frame(x = 443674, y = 4636999))

intensity_point
```

The intercept of the estimated log intensity function is UTM coordinate -31,400

The value of the intensity function at (x = 443674, y = 4636999) is $7.705 * 10^{-7}$.


