---
title: "Time Series Introduction Question 2"
output: html_document
date: "2023-06-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(mvtnorm)
library(tidyverse)

#Creates a data.frame object, the easy structure to use for ggploting
tennisCourt = data.frame(x1 = c(0,4.5,18,31.5,36,0,4.5,4.5,0,-2),
                         x2 = c(0,4.5,18,31.5,36,36,31.5,31.5,36,38),
                         y1 = c(-39,-39,-21,-39,-39,39,21,-21,-39,0), 
                         y2 = c(39,39,21,39,39,39,21,-21,-39,0),
                         width = c(rep(1,9),3))

ggTennis = ggplot(tennisCourt) + 
  geom_segment(aes(x = x1,y = y1,xend = x2,yend = y2),size = tennisCourt$width) + 
  labs(x = "Court Width",y = 'Court Depth',
       title = 'Tennis Court')


```
## Part (a)
```{r }
mu_vect = c(29, 16)
sigma = matrix(c(4,4,4,16), nrow = 2, ncol = 2, byrow = TRUE)
ball_data = data.frame(rmvnorm(5000, mu_vect, sigma))

ggTennis_with_ball_hits = ggTennis + 
  geom_point(data = ball_data,aes(x = X1, y = X2),color = 'firebrick')

ggTennis_with_ball_hits
```

## Part (b)
```{r }
pmvnorm(lower = c(18,0), upper = c(31.5,21), mean = mu_vect, sigma = sigma)
```

**Answer: From this analysis, the theoretical probability of a legal serve is 0.8237.**

To approximate the probability, we will run 5,000 realizations of the multivariate normal and count how many fall into the interval of interest where $18 \leq X_1 \leq 31.5$ and $0 \leq X_2 \leq 21$. We then divide this number by 5,000 to yield the percentage of realizations that fell within the intervals. This should be very close to our theoretical answer. We do this as follows:

```{r}
data_in_interval = ball_data %>% 
  filter(
    between(ball_data$X1, 18, 31.5),
    between(ball_data$X2, 0, 21)
    )

nrow(data_in_interval) / nrow(ball_data)
```

**Answer: Thus, our approximation yields `r nrow(data_in_interval)` realizations in the interval bounds, which we divide by `r nrow(ball_data)` to yield an approximated probability of `r nrow(data_in_interval) / nrow(ball_data)`**

## Part (c)
```{r}
x1 = 30.5

mu_condit = mu_vect[2] + sigma[1,2]%*%solve(sigma[1,1])*(x1 - mu_vect[1])

sigma_condit = sigma[2,2] - sigma[1,2]%*%solve(sigma[1,1])%*%sigma[2,1]
```

**Answer: Then the conditional distribution $X_2|X_1=30.5 \sim N($ `r mu_condit`, `r sigma_condit` $)$**

Given that $X_1 = 30.5$, the probability that serves are legal is found as follows:
```{r}
condit_prob_legal = pnorm(21, mean = mu_condit, sd = sqrt(sigma_condit)) - pnorm(0, mean = mu_condit, sd = sqrt(sigma_condit))

condit_prob_legal
```
**Answer: So the probability that a serve is legal given that $X_1 = 30.5$ is `r condit_prob_legal`**

## Part (d)
```{r}
new_data = data.frame(rnorm(500, mean = mu_condit, sd = sqrt(sigma_condit)))

colnames(new_data) = 'X2'

new_ggTennis = ggTennis + 
  geom_point(data = ball_data,aes(x = X1, y = X2),color = 'firebrick') +
  geom_point(data = new_data,aes(x = 30.5, y = X2),
             color = '#56B4E9',
             position = position_jitter(
               width = 0.6,  # amount of jitter in horizontal direction
               height = 0     # amount of jitter in vertical direction (0 = none)
               ),
             size = 0.3
    )
  

new_ggTennis
```

**Answer: The realizations of the conditional distribution form a normal curve in the $X_2$ direction, with all plotted points being on the vertical line $X_1 = 30.5$. The values generated from the conditional distribution are centered slightly higher than the original marginal distribution for $X_2$ (the conditional mean is 17.5, while the original marginal mean is 16). Additionally, the values in the conditional distribution are less spread out compared to the original marginal distribution for $X_2$ (the conditional variance is 12, while the original marginal variance is 16).**


